Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory

### 2021-01-12 papers/thesis-kirstie/thesis.tex
edits

There can be many different number density profiles for a given number
density. The number density is the number of atoms per volume, whereas
the number density profile describes how much "atom" will be found at various 
points in the cell. 
It can be thought of as an average over a large collection of actual cells
of various possible spatial distributions of the atoms within a cell, or
more precisely, as an ensemble average over all possible microstates, 
each of which consists of a unique set of positions for the atoms in 
that occupy the cell. Mathematically, the microstates are represented 
by a set of delta functions.
\begin{align}
    n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~
\end{align} 
By varying the width of the Gaussian distributions, a number of different 
number density profiles can be generated for a given number density.
It is important to note that the number density does not change when 
introducing a number density profile.  
There are still the same number of atoms over some fixed volume. 
Taking the average of the number density profile over some volume 
gives the number density. 
\begin{align}
    n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}
\end{align}

@@

Besides changing the width of the Gaussians, there is another way to 
generate different number density profiles for a given number density.
It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping the overall number density the same by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure~\ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians each of which is now one-eighth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eighth of an 
atom in each small cell as shown in Figure~\ref{fig:SameStatPic}.  
The number of particles within a primitive cell is then given by  
\begin{align}
    N= (1-f_v){N_{no.vacancies}}
\end{align}

@@

Once the Helmholtz free energy for both the crystal and fluid states of 
the fluid are found for a given temperature and density, they can be 
used to construct phase diagrams.
Outside the region of coexistence which exists between the fluid and 
crystal states, the fluid will be completely in a crystalline state if the 
crystal free energy is lower than the fluid free energy, and it will be 
completely in a fluid state if the fluid free energy is lower than 
the crystal free energy. 
We determine the actual number densities of the crystal and fluid states
at the boundaries of the coexistence region for a given temperature, 
and construct phase diagrams as described in Section 2.2.

\section{Our Functional for the Excess Helmholtz free energy}
In order to apply classical Density Functional Theory, an expression 
for the Helmholtz free energy as a functional of a number density profile 
must be found. Computing the ideal portion of the 
Helmholtz free energy is straightforward, but computing the excess portion 
of the Helmholtz free energy is difficult since it depends on the interatomic 
potential energy $V(r)$ that exists between each sphere with all the other spheres, 
as is shown by Equations~\ref{eq:mayerfunction} and \ref{eq:Fexcess-simplified}.
Soft Fundamental Measure Theory is a method
developed to form an expression for the excess Helmholtz free energy without
explicitly taking into account the interaction of each sphere with all the others, 
but it must be tailored to fit a particular fluid\cite{schmidt1999density}.

Our SFMT functional is tailored to fit a WCA fluid, and 
is written as 
\begin{align}\label{eq:Fexfunctional}
  F_{excess}[n(\vec{r})]=k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}
\end{align}
%where 
\begin{align}
\Phi_1 &= -n_{0}\ln(1-n_{3}) \\
\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align} 
The weighted densities $n_0, n_1, n_2,$ and $n_3$ are found from
\begin{align}\label{eq:numdenprofile}
	n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}
\end{align}
where the number density profile $n(\vec r)$ within the integral is given 
by Gaussian functions centered at each lattice point.
There are four scalar weight functions given by
\begin{align}\label{eq:weights}
  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
  w_2(r) &=-\frac{\partial{w_3(r)}}{\partial{r}} \\
  &= \frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}  \\
  w_3(r) &=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]  
\end{align}
two vector weight functions given by 
\begin{equation}\label{eq:w_v2}{  \vec{w_{2}}=w_{2}\frac{\vec{r}}{r}  }\end{equation}
\begin{equation}\label{eq:w_v1}{  \vec{w_{1}}=w_{1}\frac{\vec{r}}{r}  }\end{equation} 
and one tensor weight function which is given by
\begin{equation}\label{eq:tensorweight-methods}\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)\end{equation}
All of the weight functions depend on the expression for $w_3(r)$ 
which in turn depends on the interatomic potential $V(r)$ that exists 
between the spheres, and which defines the particular fluid being modeled.

The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
Schmidt's proposed error function model~\cite{schmidt2000fluid} designed 
to model a steep, soft potential in which he set
\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
subject to the boundary conditions $w_3(0)\approx{1}$ 
and $w_3(\infty)=0$. 
Our functional is  created by replacing Schmidt's parameters $\sigma$ 
and $a$ with our temperature dependent parameters $\alpha$ and $\Xi$.
The two parameters $\alpha$ and $\Xi$ make the temperature-dependent error 
function potential $V_{erf}(r)$ approximate, in the region of interest, 
the temperature-independent WCA potential which is difficult 
to implement computationally. 
%$\Xi(T)$ is determined by setting the second virial coefficient $B_2$ of 
%the error function potential $V_{erf}(r)$ equal to that of the
%WCA potential, $B_{2erf}(T)=B_{2wca}(T)$ for a given temperature.  
The error function potential emerges from $w_{3}(r)$, and the relationships 
in Equations~\ref{w2_w3_relation} and~\ref{mayerder=w2convolution}.
Using Equation~\ref{w2_w3_relation} repeated here,
\begin{displaymath}
     w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
\end{displaymath}
and solving for $w_2(r)$
\begin{align}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{align} 
we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} 
Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
\begin{displaymath}
    \frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}
\end{displaymath}
gives
\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
which can be integrated to obtain the Mayer function
\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
The Mayer function for the error function potential $V_{erf}(r)$ is thus 
given by
\begin{align}  \label{eq:VerfMayerfunction}
     f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]
\end{align} 
Combining this result with the definition of the Mayer function 
\begin{align}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{align} 
gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
\begin{equation} \label{eq:Verf}
	V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
\end{equation}

@@

Figure \ref{fig:weight_functions} shows a one dimensional view of the 
weight functions $w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance 
along the z-axis. In three dimensions, the weighting functions 
$w_0$, $w_1$, $w_2$ are spherical shells centered at $\vec{r}$, 
and $w_3$ is a solid sphere with an edge that tapers off.
$\Xi$ is related to the width of the weight functions which 
become thicker, or taper off longer as the temperature increases. 
$\alpha$ is related to the radius of the shell, or sphere, and decreases 
as the temperature increases.

\section{Computing the Fluid and Crystal Helmholtz free energies}

The fluid ideal Helmholtz free energy is computed analytically 
using Equation~\ref{eq:Fideal} 
for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
The crystal ideal Helmholtz free energy is computed computationally 
from Equation~\ref{eq:Fideal-n(r)} with the varying number density profile given by
\begin{align}
  n(|\vec{r}-\vec{R_j}|)=\sum_j{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_j}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}
\end{align}
where $\vec R_j$ is the position vector of the jth lattice site.
The fluid excess Helmholtz free energy is computed analytically  
using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
with a uniform number density profile, n that does not depend on $\vec r$.

@@

The crystal excess Helmholtz free energy is computed  
using SFMT with a number density profile 
that consists of Gaussian distributions 
centered at lattice sites specified by position vectors $\vec R$. 
The weight functions $w_i(|\vec r - \vec r~'|)$ are centered at $\vec r'$,
as shown for $w_2$ in Figure~\ref{fig:GaussandW2_actual}a. 
But because the weight function is a function of the magnitude of the 
difference between $\vec r$ and $\vec r~'$, it can be thought of as 
centered at $\vec r$, as shown in Figure~\ref{fig:GaussandW2_actual}b.
Keeping the weight function centered at $\vec r$, 
Monte-Carlo integration can be used to compute the weighted densities $n_i(\vec r)$
by randomly generating vectors
$\vec r~' -\vec R$ according to a Gaussian distribution. 
Only lengths for $\vec r~'- \vec R$ that are within an inclusion radius
about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
will be taken into account as Gaussians 
farther away will not contribute much.
The equation for the contribution to the weighted density from a 
Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
\begin{equation}{n_i(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_i(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
and is implemented computationally in the form
\begin{equation}{n_i(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_i(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
which can be written as
\begin{equation}{n_i(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}~'|)}\end{equation} 
NumPoints is the number of vectors $\vec r~' - \vec R_j$ randomly 
generated about the jth lattice site, and $n(\vec r~' - \vec R_j)$
is equal to a probability density which when multiplied by dV gives 
the probability $P_{r'}$ at which the value of $w_i(|\vec{r}-\vec{r}~'|)$ 
will occur. Using the relation
\begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}~'|)$:
\begin{equation}{n_i(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
This process is repeated to get the contributions to the weighted density 
$n_i(\vec r)$ from the Gaussians around each lattice point which begin to 
merge together as the width of the Gaussians increases.
All the contributions are then added together 
to get the final value of the weighted density at some point $\vec r$.

@@

Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature 
$T^*=\frac{kT}{\epsilon}$ is plotted against the reduced number density 
$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
It is seen that in general, a fluid exists at lower number densities, 
and a solid exists at higher number densities. 
For a given temperature, it is seen that as the density of the 
fluid increases, a maximum number density is reached for the fluid. 
Looking at the coexistence region as a whole for all temperatures, 
it is seen to extend over a narrow range of number densities which 
gradually shifts towards higher densities as the temperature increases. 
This is expected since at higher temperatures the fluid would have to 
be under greater compression before forming a solid.

@@

Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure 
$p^*=\frac{P\sigma^3}{\epsilon}$ plotted against the reduced temperature 
for temperatures of kT = 0.5 to 3.0. 
The black line marks the phase boundary between the solid and fluid regions. 
It is seen that the higher the temperature, the higher the pressure must be
to undergo a phase transition from a fluid to a solid, as would be expected. 
Ideally, the slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) 
is given by
\begin{align} 
    \frac{dP}{dT}=\frac{S_l-S_s}{V_l-V_s}
\end{align}
Since the entropy and the volume of the fluid is greater than the entropy
and the volume of the solid, the slope should be positive.

Comparing the slope of the phase boundary predicted by cDFT to that
from Monte-Carlo simulations shown by the white line, it is seen that
the cDFT transition temperatures are consistently higher, and the 
discrepancy grows as the temperature increases. For example, at a 
temperature of kT=1, the phase transition reduced pressure predicted by 
cDFT is 15.24, whereas the Monte-Carlo simulation predicted a value of 9.99,
and at a temperature of kT=3, cDFT gives a value of 77.27 whereas the 
Monte-Carlo simulation predicts 43.80.
This may indicate that either the difference between the fluid 
and solid phase transition entropies predicted by cDFT is too high, or 
the difference in the volumes is too low.

@@

\chapter{Conclusion}
We were successful in predicting the freezing of a Weeks-Chandler-Anderson 
fluid using classical Density Functional Theory. To construct a Helmholtz 
free energy functional $F[n(r)]$, we used SFMT with Schmidt's error 
function model, and defined two temperature dependent parameters, $\alpha$ 
and $\Xi$, to model a WCA fluid. 
$\alpha$ was found by setting the error function potential equal to the 
WCA potential at $r=\alpha$, and $\Xi$ was found for each temperature by 
setting the second virial coefficients for the error function potential 
and the WCA potential equal to each other, $B_{2erf}(T)=B_{2wca}(T)$.
To implement cDFT we used Gaussian distribution functions centered at 
lattice points to represent the number density profile which was then varied
by changing the width of the Gaussians, and the fraction of vacant lattice
sites until the Helmholtz free energy was minimized. Once the equilibrium 
Helmholtz free energy for the fluid and solid states were found for a given
temperature and number density, 
we used this data 
to find the number densities of the fluid and the solid at the phase transition.
Using this information, we generated a temperature-density 
phase diagram that yielded coexistence regions that approximately 
matched the Monte-Carlo
% and molecular simulation 
results, though shifted a bit toward higher densities. 
%for values of kT=0.5 to 3. 
%kT=0.5 to 38
We also generated a pressure-temperature
phase diagram, but the slope was notably steeper than that
predicted from Monte-Carlo simulations, giving transition 
temperatures and pressures that were too high.

>>>>

There can be many different number density profiles for a given number
density. The number density is the number of atoms per volume, whereas
the number density profile describes how much "atom" will be found at various 
points in the cell. 
It can be thought of as an average over a large collection of actual cells
with various possible spatial distributions of the atoms within the cell, or
more precisely, as an ensemble average over all possible microstates, 
each of which consists of a unique set of positions for the atoms 
that occupy the cell. Mathematically, the microstates are represented 
by a set of delta functions.
\begin{align}
    n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~
\end{align} 
By varying the width of the Gaussian distributions, a number of different 
number density profiles can be generated for a given number density.
It is important to note that the number density does not change when 
introducing a number density profile.  
There are still the same number of atoms over some fixed volume. 
Taking the average of the number density profile over some volume 
gives the number density. 
\begin{align}
    n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}
\end{align}

@@

There is another way to generate different number density profiles for 
a given number density besides changing the width of the Gaussians. 
It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping the overall number density the same by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure~\ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians each of which is now one-eighth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eighth of an 
atom in each small cell as shown in Figure~\ref{fig:SameStatPic}.  
The number of particles within a primitive cell is then given by  
\begin{align}
    N= (1-f_v){N_{no.vacancies}}
\end{align}

@@

Once the Helmholtz free energy for both the crystal and fluid 
are found for a given temperature and density, they can be 
used to construct phase diagrams.
Outside the region of coexistence, a purely crystalline state will exist 
if the crystal free energy is lower than the fluid free energy, and 
a purely fluid state will exist if the fluid free energy is lower than 
the crystal free energy. 
After the actual number densities of the crystal and fluid states
at the boundaries of the coexistence region are determined for a given 
temperature, phase diagrams can be constructed as described in Section 2.2.

\section{Our Functional for the Excess Helmholtz free energy}
In order to apply classical Density Functional Theory, an expression 
for the Helmholtz free energy as a functional of a number density profile 
must be found. Computing the ideal portion of the 
Helmholtz free energy is straightforward, but computing the excess portion 
of the Helmholtz free energy is difficult since it depends on the interatomic 
potential energy $V(r)$ that exists between each sphere and all the other spheres, 
as is shown by Equations~\ref{eq:mayerfunction} and \ref{eq:Fexcess-simplified}.
Soft Fundamental Measure Theory is a method
developed to form an expression for the excess Helmholtz free energy without
explicitly taking into account the interaction of each sphere with all the others, 
but it must be tailored to fit a particular fluid\cite{schmidt1999density}.

Our SFMT functional is tailored to fit a WCA fluid, and 
is written as 
\begin{align}\label{eq:Fexfunctional}
  F_{excess}[n(\vec{r})]=k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}
\end{align}
%where 
\begin{align}
\Phi_1 &= -n_{0}\ln(1-n_{3}) \\
\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align} 
The weighted densities $n_0, n_1, n_2, n_3, \vec {n}_1, \vec {n}_2$ and $\overleftrightarrow{n}_{m2}$ are found from
\begin{align}\label{eq:numdenprofile}
	n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(|\vec{r}-\vec{r'}|)d{\vec{r'}}}
\end{align}
where the number density profile $n(\vec r)$ within the integral is given 
by Gaussian functions centered at each lattice point.
There are four scalar weight functions given by
\begin{align}\label{eq:weights}
  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
  w_2(r) &=-\frac{\partial{w_3(r)}}{\partial{r}} \\
  &= \frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}  \\
  w_3(r) &=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]  
\end{align}
two vector weight functions given by 
\begin{equation}\label{eq:w_v2}{  \vec{w_{2}}=w_{2}\frac{\vec{r}}{r}  }\end{equation}
\begin{equation}\label{eq:w_v1}{  \vec{w_{1}}=w_{1}\frac{\vec{r}}{r}  }\end{equation} 
and one tensor weight function which is given by
\begin{equation}\label{eq:tensorweight-methods}\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)\end{equation}
All of the weight functions depend on the expression for $w_3(r)$ 
which in turn depends on the interatomic potential $V(r)$ that exists 
between the spheres, and which defines the particular fluid being modeled.

The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
Schmidt's proposed error function model~\cite{schmidt2000fluid} designed 
to model a steep, soft potential in which he set
\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
subject to the boundary conditions $w_3(0)\approx{1}$ 
and $w_3(\infty)=0$. 
Our functional is  created by replacing Schmidt's parameters $\sigma$ 
and $a$ with our temperature dependent parameters $\alpha$ and $\Xi$.
%The two parameters $\alpha$ and $\Xi$ make the temperature-dependent error 
%function potential $V_{erf}(r)$ approximate, in the region of interest, 
%the temperature-independent WCA potential which is difficult 
%to implement computationally. 
The two parameters $\alpha$ and $\Xi$ are designed to make the functional
produce results which approximate, in the region of interest, that of  
a WCA fluid using the temperature-dependent error function potential $V_{erf}(r)$ 
instead of the temperature-independent WCA potential which is difficult 
to implement computationally. 
%$\Xi(T)$ is determined by setting the second virial coefficient $B_2$ of 
%the error function potential $V_{erf}(r)$ equal to that of the
%WCA potential, $B_{2erf}(T)=B_{2wca}(T)$ for a given temperature.  
The error function potential emerges from $w_{3}(r)$, and the relationships 
in Equations~\ref{w2_w3_relation} and~\ref{mayerder=w2convolution}.
Using Equation~\ref{w2_w3_relation}(repeated below),
\begin{displaymath}
     w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
\end{displaymath}
and solving for $w_2(r)$
\begin{align}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{align} 
we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} 
Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
\begin{displaymath}
    \frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}
\end{displaymath}
gives
\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
which can be integrated to obtain the Mayer function
\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
The Mayer function for the error function potential $V_{erf}(r)$ is thus 
given by
\begin{align}  \label{eq:VerfMayerfunction}
     f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]
\end{align} 
Combining this result with the definition of the Mayer function 
\begin{align}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{align} 
gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
\begin{equation} \label{eq:Verf}
	V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
\end{equation}

@@

Figure \ref{fig:weight_functions} shows a one dimensional view of the 
weight functions $w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance 
along the z-axis. In three dimensions, the weighting functions 
$w_0$, $w_1$, $w_2$ are spherical shells, 
and $w_3$ is a solid sphere with an edge that tapers off.
$\Xi$ is related to the width of the weight functions which 
become thicker, or taper off longer as the temperature increases, and 
$\alpha$ is related to the radius of the shell, or sphere, and decreases 
as the temperature increases.

\section{Computing the Fluid and Crystal Helmholtz free energies}
The Helmholtz free energy is composed of two parts: an ideal Helmholtz free energy 
and an excess Helmholtz free energy that are added together to get the total
Helmholtz free energy, $F=F_{ideal} + F_{excess}$. 
And so the ideal and excess Helmholtz free energies
must be computed for both the fluid and the crystal states.
The fluid ideal Helmholtz free energy is computed analytically 
using Equation~\ref{eq:Fideal} for a given homogeneous reduced 
(or dimensionless) number density $n^*$ and temperature $t^*$. 
The crystal ideal Helmholtz free energy is computed computationally 
from Equation~\ref{eq:Fideal-n(r)} with the varying number density profile given by
\begin{align}
  n(|\vec{r}-\vec{R_j}|)=\sum_j{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_j}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}
\end{align}
where $\vec R_j$ is the position vector of the jth lattice site.

@@

The fluid excess Helmholtz free energy is computed analytically  
using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
with a uniform number density profile, n that does not depend on $\vec r$. 
The crystal excess Helmholtz free energy is computed  
using SFMT with a number density profile 
that consists of Gaussian distributions 
centered at lattice sites specified by position vectors $\vec R$. 
The weight functions $w_\alpha(|\vec r - \vec r~'|)$ are centered at $\vec r'$,
as shown for $w_2$ in Figure~\ref{fig:GaussandW2_actual}a. 
But because the weight function is a function of the magnitude of the 
difference between $\vec r$ and $\vec r~'$, it can be thought of as 
centered at $\vec r$, as shown in Figure~\ref{fig:GaussandW2_actual}b.
Keeping the weight function centered at $\vec r$, 
Monte-Carlo integration can be used to compute the weighted densities $n_\alpha(\vec r)$
by randomly generating vectors
$\vec r~' -\vec R$ according to a Gaussian distribution. 
Only lengths for $\vec r~'- \vec R$ that are within an inclusion radius
about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
will be taken into account as Gaussians 
farther away will not contribute much.
The equation for the contribution to the weighted density from a 
Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
\begin{equation}{n_\alpha(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
and is implemented computationally in the form
\begin{equation}{n_\alpha(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
which can be written as
\begin{equation}{n_\alpha(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_\alpha(|\vec{r}-\vec{r}~'|)}\end{equation} 
NumPoints is the number of vectors $\vec r~' - \vec R_j$ randomly 
generated about the jth lattice site, and $n(\vec r~' - \vec R_j)$
is equal to a probability density which when multiplied by dV gives 
the probability $P_{r'}$ at which the value of $w_\alpha(|\vec{r}-\vec{r}~'|)$ 
will occur. Using the relation
\begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
$n_\alpha(\vec{r})$ can be expressed as the average value of $(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)$:
\begin{equation}{n_\alpha(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
This process is repeated to get the contributions to the weighted density 
$n_\alpha(\vec r)$ from the Gaussians around each lattice point which begin to 
merge together as the width of the Gaussians increases.
All the contributions are then added together 
to get the final value of the weighted density at some point $\vec r$.

@@

Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
and \ref{fig:Phase_Diagram_P_vs_T}. 
In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature 
$T^*=\frac{kT}{\epsilon}$ is plotted against the reduced number density 
$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
In all cases $\sigma=1$ and $\epsilon=1$. 
It is seen that in general, a fluid exists at lower number densities, 
and a solid exists at higher number densities. 
For a given temperature, it is seen that as the density increases, 
a maximum number density is reached for the fluid. 
Looking at the coexistence region as a whole for all temperatures, 
it is seen to extend over a narrow range of number densities which 
gradually shifts towards higher densities as the temperature increases. 
This is expected since at higher temperatures the fluid would have to 
be under greater compression before forming a solid.

@@

Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure 
$p^*=\frac{P\sigma^3}{\epsilon}$ plotted against the reduced temperature 
for temperatures of kT = 0.5 to 3.0. 
The black line marks the phase boundary between the solid and fluid regions. 
It is seen that the higher the temperature, the higher the pressure must be
to undergo a phase transition from a fluid to a solid, as would be expected. 
Ideally, the slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) 
is given by
\begin{align} 
    \frac{dP}{dT}=\frac{S_f-S_s}{V_f-V_s}
\end{align}
Since the entropy and the volume of the fluid is greater than the entropy
and the volume of the solid, the slope should be positive.

Comparing the slope of the phase boundary predicted by cDFT to that
from Monte-Carlo simulations shown by the white line, it is seen that
the phase transition pressures and temperatures predicted by cDFT 
are consistently higher, and the 
discrepancy grows as the temperature increases. For example, at a 
temperature of kT=1, the phase transition reduced pressure predicted by 
cDFT is 15.24, whereas the Monte-Carlo simulation predicted a reduced 
pressure of 9.99,
and at a temperature of kT=3, cDFT gives a value of 77.27 whereas the 
Monte-Carlo simulation predicts 43.80.
This may indicate that either the difference between the fluid 
and solid phase transition entropies predicted by cDFT is too high, or 
the difference in the volumes is too low.

@@

\chapter{Conclusion}
We were successful in predicting the freezing of a Weeks-Chandler-Anderson 
fluid using classical Density Functional Theory. To construct a Helmholtz 
free energy functional $F[n(r)]$, we used SFMT with Schmidt's error 
function model, and defined two temperature dependent parameters, $\alpha$ 
and $\Xi$, to model a WCA fluid. 
$\alpha$ was found by setting the error function potential equal to the 
WCA potential at $r=\alpha$, and $\Xi$ was found for each temperature by 
setting the second virial coefficients for the error function potential 
and the WCA potential equal to each other, $B_{2erf}(T)=B_{2wca}(T)$ 
at each temperature.
To implement cDFT we used Gaussian distribution functions centered at 
lattice points to represent the number density profile which was then varied
by changing the width of the Gaussians, and the fraction of vacant lattice
sites until the Helmholtz free energy was minimized. Once the equilibrium 
Helmholtz free energy for the fluid and solid states were found for a given
temperature and number density, 
we used this data 
to find the number densities of the fluid and the solid at the phase transition.
Using this information, we generated a temperature-density 
phase diagram that yielded coexistence regions that approximately 
matched the Monte-Carlo
% and molecular simulation 
results, though shifted a bit toward higher densities. 
%for values of kT=0.5 to 3. 
%kT=0.5 to 38
We also generated a pressure-temperature
phase diagram, but the slope was notably steeper than that
predicted from Monte-Carlo simulations, giving transition 
temperatures and pressures that were too high.
